---
title: "Exercise 09"
output: html_document
date: "2024-03-18"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Step 1 
- Using the {tidyverse} read_csv() function, load the “Street_et_al_2017.csv” dataset from this URL as a “tibble” named d.
- Do a quick exploratory data analysis where you generate the five-number summary (median, minimum and maximum and 1st and 3rd quartile values), plus mean and standard deviation, for each quantitative variable.

```{r echo=TRUE, message=FALSE}
# Set up libraries
library(tidyverse)
library(skimr)
```

```{r, 1}
f <- "https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/Street_et_al_2017.csv"
d <- read_csv(f, col_names = TRUE)

# Quick glance
head(d)

# From {skimr}; provides n_missing, mean, sd, and five number summary (min, max, Q1, median, Q3)
skim(d)

# Make it a tibble 
# (tibble_d <- skim(d) 
#  tibble::as_tibble())

# You can also easily use group_by function to ...group...by... variables
# d %>%
#  dplyr::group_by(Species) %>%
#  skim()

```

### Step 2 
- From this dataset, plot brain size (ECV) as a function of social group size (Group_size), longevity (Longevity), juvenile period length (Weaning), and reproductive lifespan (Repro_lifespan).

```{r, 2}
summary(cars)
```

### Step 3
-Derive by hand the ordinary least squares regression coefficients 

```{r, 3}

```

### Step 4


```{r, 4}

```


### Step 5


```{r, 5}

```

### Step 6


```{r, 6}

```

### Step 7


```{r, 7}

```

### Step 8 

```{r, 8}


```
